Easy as Pi, Volume II: On the Goodput and Energy of Network Coding Schemes in the Raspberry Pi

### 2016-08-16 reviewer_reply.txt

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 1:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,
Thank you very much for your valuable feedback. We have listed your comments below and described
our corrections to hopefully solve the indicated issues.

- We have corrected all the indicated and other found typos by proof-reading the manuscript. We
have also replaced the abbreviations "l.d." and "l.i." to "linearly dependent" and "linearly
independent" as suggested to enhance clarity in the notation.

- Regarding the definition of "g", We have defined the parameter "g" as the total number of
original packets to be encoded once at beginning of Section 2.

- For the notation [1,g] to denote the set of integer number between 1 and "g", we have modified
all its appearances to [1,2,...,g] to indicate the integer nature of this set in a more standard
fashion. Also, in our
definitions the notation

- For the notation $\mathbb{N}$ for the random number of packets to be combined in SRLNC, we have
modified all its appearances to $\mathbb{M}$ to avoid any confusion with the notation for the set of
positive numbers.

@@

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 2:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,
Thank you very much for your valuable feedback. We have listed your comments below and described
our corrections or observations to hopefully solve the indicated issues.

@@

- The proposed system is not well described

We have rewritten the parts that indicate a proposed system. All our measurements are performed
within a single Raspberry Pi. We only replace the traditional power supply with an Agilent Power
Supply device to monitor and log the power consumption.

- Finally and in a word, the motivation of this paper is unclear. The main question is "why
Raspberry pi in network coding?". What is the advantage of using Raspberry pi?

We consider that the Raspberry Pi models 1 and 2 are good representatives of a low-cost computer
considered as IoT devices in terms of performance and capabilities. We now elaborate on this idea
in the introduction by providing various use cases of IoT + Raspi. At the end of the description of
the use cases of IoT + Raspi, we specifically mention that despite this advances, data is still
being conveyed in a conventional way that will hit a performance roof with new services demands.
This is why we motivate about network coding, e.g. since the usefulness of network coding has been
verified in the area of meshed networks in prior works. Meshed networks will be the scenario where
upcoming IoT deployments will mostly occur.

Another two important reasons for using the Raspberry Pi are: First, its large user community in
contrast to many other low-cost devices. This allows a system developer to use standard tools
and / or based on either Windows or Linux, reducing the development time. Second, the Raspberry
Pi models 1 and 2, have one and four cores respectively and they support hardware acceleration
through the NEON instruction set in ARM that is mandatory to reach the full potential of network
coding where we have included the proper details. We have updated the introduction to elaborate
on all these subjects to show why using network coding in Raspberry Pi is interesting.

- They only consider q=2 and 28 , however, as it is mentioned in "Chou, Philip A., Yunnan Wu, and
Kamal Jain. "Practical network coding." (2003).", based on security aspect, in practice, wey
should consider q=28 and 216. By this assumption, they should show and compare the results with
q=216 too.

We have made a reference to two papers by Achuthan et al. (i.e. "Lean and Mean" [23] and "Sharing
the Pi" [25]). These two references include measurements for both q=2^8 and q=2^16 and indicate
a performance degradation of approximately 10X for going from q=2^8 to q=2^16. The underlying
reason is that using Galois fields higher than 2^8 provide no perceivable benefit in terms of the
decoding probability while increasing in complexity and processing overhead due to the coding
coefficients. At the end, this significantly reduces the performance of the code. To verify this,
we also ran a few measurements ourself that indicated a similar performance gap between the two
field types. Another reason to not consider q=2^16 is that it is not a standard field size. q=2^8
is the standard today (comparable with Reed-Solomon codes for example) and q=2 is beneficial in
terms of performance in low-computing power devices. Finally, for security purposes, there exists
much secure ways than only using q=2^16.

@@

- The list of references neglects some of the most relevant papers. Moreover, the authors have
cited to their previous papers (i.e., [6,7,8,10,11,12,13,17,21,27,30])!!! It seems they want to
increase citation of their papers!

We have updated the list of references to provide a richer balance in the State-of-the-Art,
relevant baseline references and our prior work.

We now have included several references to previous State-of-the-Art and references related
to use cases for IoT employing the Raspberry Pi [1-9]. Also we have included missing
references and grouped them with the baseline of network coding [10-16,26].

We link to our prior work as this paper evaluation is based on research lines started
in those papers. These prior works report our findings in the area of network coding for
wireless mesh networks and network coding for the Raspberry Pi. To be more specific: Theoretical
gains and practical implementations of network coding for wireless
meshed networks are reported in [17,19,21,22]. The Kodo library, our basic building block for the coding operations is [18]. Basic measurements + observed effects of Random
Linear Network Coding for commercial devices are found in [23]. A initial study of RLNC for
the Raspberry Pi is available in [25]. Finally, our work with Sparse and Tunable Sparse Network
Coding is described in [28-31].

At the end, our references consist of less than 27% of the references in the study. We hope
that this solves the raised issue.

- The presentation of the paper is not very good. The authors did not care very much about the
structure, the language (i.e., page 2 line 42 "Furthermore, in [7], the Kodo library was
introduce." Should be introduced....)

### 2016-08-16 introduction.tex
Corrected version. Update review reply. Update readme.rst

Compared to other broadly used coding schemes, such as \ac{LDPC}
codes~\cite{gallager1962low} or Reed-Solomon
codes~\cite{reed1960polynomial}, network coding is a technology that
has been implemented in real systems since the early years of its
conception. Authors in~\cite{katti2008xors} showed the
performance of the COPE protocol for the two-way relay channel
in a wireless network, which relied on minimalistic coding and obtaining
gains over a forwarding scheme. Later, the work in~\cite{pedersen2008implementation} used commercially available
Symbian OS mobile phones to implement network coding in a
\ac{D2D} cooperation-based application. Furthermore,
in~\cite{kodo2011pedersen}, the Kodo library was introduced. Kodo is
a C++11 network coding library intended to make network coding
basic functionalities easily available for both the research community
and commercial entities. Based on Kodo, the CATWOMAN
protocol~\cite{hundeboll2012catwoman} is implemented on top of the
BATMAN protocol~\cite{johnson2008simple} for WiFi multi-hop meshed
networks. It uses some of the intuition from COPE, but it is deployed
within the X topology with overhearing links. Its source code is available
as open source in the Linux kernel. Moreover, many other successful
implementations have been tested on real world systems such as
found in~\cite{pahlevani2013playncool,krigslund2013core,
paramanathan2013leanandmean,seferoglu2011i2nc}. For the Raspberry Pi device, an evaluation
of the performance of a decentralized network code for distributed systems, \ac{RLNC},
\cite{ho2006random}, can be found in \cite{paramanathan2014sharing}.
However, this evaluation focused particularly on observing the achievable
speeds only for \ac{RLNC} with different configurations of
the code parameters. In this previous work, a relevant practical aspect that
was not evaluated was the use of hardware acceleration through \ac{SIMD} or
the multi-core capabilities of more advanced \ac{Raspi} models. These features are
becoming more frequent in new processors to largely increase their computational
power for the upcoming demand. In this sense, the Raspberry Pi posses an \ac{ARM}
architecture that could be multi-core as mentioned and also exploits the \ac{SIMD}
feature with the optimized NEON instruction set, but still to the best of our knowledge
there has been no documentation in the literature about these capabilities.

>>>>

Compared to other broadly used coding schemes, such as \ac{LDPC}
codes~\cite{gallager1962low} or Reed-Solomon
codes~\cite{reed1960polynomial}, network coding is a technology that
has been studied and implemented in real systems since the early years of
its conception. A decentralized network code that has been proven
to achieve the multicast capacity with very high probability is
\ac{RLNC} \cite{ho2006random}. Later, a work observing the
benefits of employing \ac{RLNC} in meshed networks is the MORE protocol
addressed in \cite{chachulski2007more}. Shortly afterwards, authors
in~\cite{katti2008xors} showed the performance of an implementation
of the COPE protocol for the two-way relay channel
in a wireless network, which relied on minimalistic coding and obtaining
gains over a forwarding scheme. Later, the work
in~\cite{pedersen2008implementation} used commercially available
Symbian OS mobile phones to implement network coding in a
\ac{D2D} cooperation-based application. Furthermore,
in~\cite{kodo2011pedersen}, the Kodo library was introduced. Kodo is
a C++11 network coding library intended to make network coding
basic functionalities easily available for both the research community
and commercial entities. Based on Kodo, the CATWOMAN
protocol~\cite{hundeboll2012catwoman} is implemented on top of the
BATMAN protocol~\cite{johnson2008simple} for WiFi multi-hop meshed
networks. It uses some of the intuition from COPE, but it is deployed
within the X topology with overhearing links. Its source code is available
as open source in the Linux kernel. Moreover, many other successful
implementations have been tested on real world systems such as
found in~\cite{pahlevani2013playncool,krigslund2013core,
paramanathan2013leanandmean,seferoglu2011i2nc}. For the Raspberry Pi device, an
evaluation of \ac{RLNC} can be found in \cite{paramanathan2014sharing}.
However, this evaluation focused particularly on observing the achievable
speeds only for \ac{RLNC} with different configurations of
the code parameters. In this previous work, a relevant practical aspect that
was not evaluated was the use of hardware acceleration through \ac{SIMD} or
the multi-core capabilities of more advanced \ac{Raspi} models. These features are
becoming more frequent in new processors to largely increase their computational
power for the upcoming demand. In this sense, the Raspberry Pi posses an \ac{ARM}
architecture that could be multi-core as mentioned and also exploits the \ac{SIMD}
feature with the optimized NEON instruction set, but still to the best of our knowledge
there has been no documentation in the literature about these capabilities.

### 2016-08-16 schemes.tex
Corrected version. Update review reply. Update readme.rst

\ac{RLNC} is an example of intra-session \ac{NC}, i.e., data symbols
from a single flow are combined with each other. In this type of
%network coding, all the original data packets $P_j, j \in [1,g]$, each
network coding, $g$ original data packets $P_j, j \in [1,2,\ldots,g]$, each
of $B$ bytes, are used to create coded packets using random linear
combinations of the original ones. In the following subsections, we
describe the basic functionalities of \ac{RLNC}.

\subsubsection{Encoding}
In \ac{RLNC}, any coded packet is a linear combination of all
the original packets. For the coding scheme, packets are seen as
algebraic entities formed as a sequence of elements from $GF(q)$,
which is a \ac{GF} of size $q$. Later, each original packet is
multiplied by a coding coefficient from $GF(q)$. The coding coefficients
are chosen uniformly at random from the \ac{GF} by the encoder. To
perform the multiplication of a packet by a coding coefficient, the
coefficient is multiplied for each of the elements in the
concatenation that composes an original packet, preserving the
concatenation. Later, all resulting packets are added within the
\ac{GF} arithmetics together to generate a coded packet. Thus, a
coded packet can be written as:
%
\begin{align} \label{eq:coded_packet}
C_i  &= \bigoplus_{j=1}^{g} v_{ij} \otimes P_j ,\ \forall i \in [1,\ldots)
\end{align}

@@

In \eqref{eq:recoded_packet}, $w_{ij}$ is the coding coefficient used
in the $i$-th recoded packet and assigned to multiply a previously
coded packet $C_j$. These coding coefficients are again uniformly and
randomly chosen from $GF(q)$. However, these $w_{ij}$'s are not
appended to the previous coding coefficients.  Instead, the system
will update the previous coefficients. Due to the linearity of the
operation, this update reduces to recombining the coding coefficients
equivalent to each original packet with the weight of the
$w_{i,j}$.Therefore, we define the local coding matrix $\textbf{W}$ in
the same way as it was made for $\textbf{V}$. Thus, the local coding
matrix can be written as:
%
\begin{align} \label{eq:local_coding_matrix}
\textbf{W} =
\left[
\begin{array}{c}
        w_1    \\ \hline
        \vdots \\ \hline
        w_g    \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
        w_{11} & \ldots & w_{1g} \\
        \vdots  & \ddots & \vdots  \\
        w_{g1} & \ldots & w_{gg} \\
\end{array}
\right].
\end{align}
%
With the definitions from \eqref{eq:coding_matrix} and
\eqref{eq:local_coding_matrix}, the recoded packets
$\textbf{R} = \left[R_1 \ldots R_g \right]^T$ are written as
$\textbf{R} = (\textbf{W} \textbf{V}) \textbf{P}$. Here, we recognize
the relationship between original and recoded packets. The resulting
coding matrix is the multiplication of matrices $\textbf{W}$ and
$\textbf{V}$. Denoting ${(\textbf{W} \textbf{V})}_{ij}$ as the element
in the $i$-th row and $j$-th column of $(\textbf{W} \textbf{V})$, this
term is the resulting coding coefficient used to create $R_i$ after
encoding the original packet $P_j$ recoding it locally in an intermediate
node. Finally, the appended coefficients for $R_i$ are
${(\textbf{W} \textbf{V})}_{ik}$ with $k \in [1,2,\ldots,g]$. By doing some
algebra, each ${(\textbf{W} \textbf{V})}_{ij}$ term can be verified to be
computed as:
%
\begin{align}
\label{eq:appended_coded_coefficients}
{(\textbf{W} \textbf{V})}_{ij} = \sum_{k=1}^{g} w_{ik} v_{kj}\ ,\ \forall i,j \in [1,g] \times [1,g].
\end{align}
%
This update procedure on the coding coefficients is carried by all the
recoders in a given network therefore allowing any decoder to compute
the original data after Gaussian elimination, regardless of the amount
of times recoding was performed and without incurring in any
additional overhead cost for signaling. Similar to the encoding
operation, any decoder that collects a set of $g$ linearly independent recoded
packets with their respective coefficients, will be able to decode the
data as mentioned before in Section \ref{sssec:decoding}.

@@

For this method, we first obtain the amount of packets
that we will combine to create a coded packet. To do so, a random
number, $\mathbb{N}$, of the original packets is used to produce a
coded packet. For our case, $\mathbb{N}$ is binomially distributed
with parameters $g$ for the number of trials and $d$ for its success
probability, e.g. $\mathbb{N} \sim Binomial(g,d)$. However,
when sampling from this distribution, the case
$\mathbb{N} = 0$ occurs with a non-zero probability. In order to handle
this special case, our implementation considers
$\mathbb{K} = \max(1,\mathbb{N})$ as the final amount of packets to be
mixed together. In this way, we always ensure that at least one packet is
encoded.

The only caveat is that the case of $\mathbb{K} = 1$ occurs slightly
more often than $\mathbb{N} = 1$ in the original distribution, but for
the considered density range in this method, this is not a significant
modification~\cite{practicalview_tsnc2015}. Then, once the
distribution for the number of packets to mix has been defined, we
sample a value $n$ from $\mathbb{N}$ and compute $k =
\max(1,n)$. Later, we create a set $\mathcal{K}$ with
cardinality $k$, e.g. $|\mathcal{K}| = k$, where the elements in
$\mathcal{K}$ are the indexes of the packets that are going to be
considered for making a coded packet. To compute the indexes
of the set $\mathcal{K}$, we do the following algorithm in pseudo-code: \\
\ \\
\begin{algorithm}[H]
 \label{alg:k_set}
 \KwData{$k$: Size of $\mathcal{K}$. $g$: Generation Size}
 \KwResult{$\mathcal{K}$: The set of non-repeated indexes}
 $\mathcal{K}$ = \{ \}\;
 \While{$|\mathcal{K}| \neq k$ }{
  $i$ = Sample from $\mathbb{U}(1,g)$\;
  \If{$i \notin \mathcal{K}$}{
   Insert $i$ in $\mathcal{K}$\;
   }
 }
 \caption{Computation of the set of indexes for packet combination in SRLNC.}
\end{algorithm}
\ \\
In Algorithm~\ref{alg:k_set}, the notation $\mathbb{U}(1,g)$ stands
for a uniform discrete random variable with limits $1$ and $g$. The
pseudo-code in Algorithm~\ref{alg:k_set} indicates that the set $\mathbb{K}$
is filled with non-repeated elements taken uniformly at random from
$[1,\ldots,g]$. Finally, once the set of indexes has been defined, a coded
packet using \ac{SRLNC} can be written as shown in
\eqref{eq:sparse_coded_packet2}:
%
\begin{align} \label{eq:sparse_coded_packet2}
    C_i  &= \bigoplus_{m \in \mathcal{K}} v_{im} \otimes P_{m},\ v_{im} \neq 0,\ \forall i \in [1,\ldots)
\end{align}

@@

Provided a desired budget $b$ and the \ac{dof} of the decoder, an encoder
can estimate the required coding density for the packets to not exceed the
budget. In our implementation, we use feedback packets to provide the
encoder an estimate of the decoder \ac{dof} at pre-defined points of the
transmission process. The points in the process occur when a decoder
obtains a given amount of \ac{dof}. In our implementation, we define $r(k)$ as the \ac{dof}s of the decoder where it has to send the $k$-th feedback in
\eqref{eq:feedback_dofs}.
%
\begin{equation}\label{eq:feedback_dofs}
    r(k) = \floor{g\cdot \left[\frac{2^k -1}{2^k}\right]},\ k \in \Big[1,\ldots,\ceil{log_2(g)}+1\Big]
\end{equation}
%
At the beginning of the transmission process, we assume that the encoder
starts with an initial coding density that has been calculated depending to
the budget. According to \eqref{eq:feedback_dofs}, a decoder sends feedback
once it has obtained (rounded down) $g/2,\ 3g/4,\ 7g/8,\ \ldots$  degrees of
freedom. The total amount of feedback for this scheme will depend on the
generation size. Still, we have an implementation that limits the amount of
feedback to be sent from the decoder in~\cite{practicalview_tsnc2015}.

>>>>

\ac{RLNC} is an example of intra-session \ac{NC}, i.e., data symbols
from a single flow are combined with each other. In this type of
%network coding, all the original data packets $P_j, j \in [1,g]$, each
network coding, $g$ original data packets also called a generation
\cite{chou2003practical}, $P_j, j \in [1,2,\ldots,g]$, each
of $B$ bytes, are used to create coded packets using random linear
combinations of the original ones. In the following subsections, we
describe the basic functionalities of \ac{RLNC}.

\subsubsection{Encoding}
In \ac{RLNC}, any coded packet is a linear combination of all
the original packets. For the coding scheme, packets are seen as
algebraic entities formed as a sequence of elements from $GF(q)$,
which is a \ac{GF} of size $q$. Later, each original packet is
multiplied by a coding coefficient from $GF(q)$. The coding coefficients
are chosen uniformly at random from the \ac{GF} by the encoder. To
perform the multiplication of a packet by a coding coefficient, the
coefficient is multiplied for each of the elements in the
concatenation that composes an original packet, preserving the
concatenation. Later, all resulting packets are added within the
\ac{GF} arithmetics together to generate a coded packet. Thus, a
coded packet can be written as:
%
\begin{align} \label{eq:coded_packet}
C_i  &= \bigoplus_{j=1}^{g} v_{ij} \otimes P_j ,\ \forall i \in [1,2,\ldots)
\end{align}

@@

In \eqref{eq:recoded_packet}, $w_{ij}$ is the coding coefficient used
in the $i$-th recoded packet and assigned to multiply a previously
coded packet $C_j$. These coding coefficients are again uniformly and
randomly chosen from $GF(q)$. However, these $w_{ij}$'s are not
appended to the previous coding coefficients.  Instead, the system
will update the previous coefficients. Due to the linearity of the
operation, this update reduces to recombining the coding coefficients
equivalent to each original packet with the weight of the
$w_{i,j}$.Therefore, we define the local coding matrix $\textbf{W}$ in
the same way as it was made for $\textbf{V}$. Thus, the local coding
matrix can be written as:
%
\begin{align} \label{eq:local_coding_matrix}
\textbf{W} =
\left[
\begin{array}{c}
        w_1    \\ \hline
        \vdots \\ \hline
        w_g    \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
        w_{11} & \ldots & w_{1g} \\
        \vdots  & \ddots & \vdots  \\
        w_{g1} & \ldots & w_{gg} \\
\end{array}
\right].
\end{align}
%
With the definitions from \eqref{eq:coding_matrix} and
\eqref{eq:local_coding_matrix}, the recoded packets
$\textbf{R} = \left[R_1 \ldots R_g \right]^T$ are written as
$\textbf{R} = (\textbf{W} \textbf{V}) \textbf{P}$. Here, we recognize
the relationship between original and recoded packets. The resulting
coding matrix is the multiplication of matrices $\textbf{W}$ and
$\textbf{V}$. Denoting ${(\textbf{W} \textbf{V})}_{ij}$ as the element
in the $i$-th row and $j$-th column of $(\textbf{W} \textbf{V})$, this
term is the resulting coding coefficient used to create $R_i$ after
encoding the original packet $P_j$ recoding it locally in an intermediate
node. Finally, the appended coefficients for $R_i$ are
${(\textbf{W} \textbf{V})}_{ik}$ with $k \in [1,2,\ldots,g]$. By doing some
algebra, each ${(\textbf{W} \textbf{V})}_{ij}$ term can be verified to be
computed as:
%
\begin{align}
\label{eq:appended_coded_coefficients}
{(\textbf{W} \textbf{V})}_{ij} = \sum_{k=1}^{g} w_{ik} v_{kj}\ ,\ \forall i,j \in [1,2,\ldots,g] \times [1,2,\ldots,g].
\end{align}
%
This update procedure on the coding coefficients is carried by all the
recoders in a given network therefore allowing any decoder to compute
the original data after Gaussian elimination, regardless of the amount
of times recoding was performed and without incurring in any
additional overhead cost for signaling. Similar to the encoding
operation, any decoder that collects a set of $g$ linearly independent recoded
packets with their respective coefficients, will be able to decode the
data as mentioned before in Section \ref{sssec:decoding}.

@@

For this method, we first obtain the amount of packets
that we will combine to create a coded packet. To do so, a random
number, $\mathbb{M}$, of the original packets is used to produce a
coded packet. For our case, $\mathbb{M}$ is binomially distributed
with parameters $g$ for the number of trials and $d$ for its success
probability, e.g. $\mathbb{M} \sim Binomial(g,d)$. However,
when sampling from this distribution, the case
$\mathbb{M} = 0$ occurs with a non-zero probability. In order to handle
this special case, our implementation considers
$\mathbb{K} = \max(1,\mathbb{M})$ as the final amount of packets to be
mixed together. In this way, we always ensure that at least one packet is
encoded.

The only caveat is that the case of $\mathbb{K} = 1$ occurs slightly
more often than $\mathbb{M} = 1$ in the original distribution, but for
the considered density range in this method, this is not a significant
modification~\cite{practicalview_tsnc2015}. Then, once the
distribution for the number of packets to mix has been defined, we
sample a value $m$ from $\mathbb{M}$ and compute $k =
\max(1,m)$. Later, we create a set $\mathcal{K}$ with
cardinality $k$, e.g. $|\mathcal{K}| = k$, where the elements in
$\mathcal{K}$ are the indexes of the packets that are going to be
considered for making a coded packet. To compute the indexes
of the set $\mathcal{K}$, we do the following algorithm in pseudo-code: \\
\ \\
\begin{algorithm}[H]
 \label{alg:k_set}
 \KwData{$k$: Size of $\mathcal{K}$. $g$: Generation Size}
 \KwResult{$\mathcal{K}$: The set of non-repeated indexes}
 $\mathcal{K}$ = \{ \}\;
 \While{$|\mathcal{K}| \neq k$ }{
  $i$ = Sample from $\mathbb{U}(1,g)$\;
  \If{$i \notin \mathcal{K}$}{
   Insert $i$ in $\mathcal{K}$\;
   }
 }
 \caption{Computation of the set of indexes for packet combination in SRLNC.}
\end{algorithm}
\ \\
In Algorithm~\ref{alg:k_set}, the notation $\mathbb{U}(1,g)$ stands
for a uniform discrete random variable with limits $1$ and $g$. The
pseudo-code in Algorithm~\ref{alg:k_set} indicates that the set $\mathbb{K}$
is filled with non-repeated elements taken uniformly at random from
the continuous interval $[1,\ldots,g]$. Finally, once the set of indexes has been
defined, a coded packet using \ac{SRLNC} can be written as shown in
\eqref{eq:sparse_coded_packet2}:
%
\begin{align} \label{eq:sparse_coded_packet2}
    C_i  &= \bigoplus_{m \in \mathcal{K}} v_{im} \otimes P_{m},\ v_{im} \neq 0,\ \forall i \in [1,\ldots)
\end{align}

@@

Provided a desired budget $b$ and the \ac{dof} of the decoder, an encoder
can estimate the required coding density for the packets to not exceed the
budget. In our implementation, we use feedback packets to provide the
encoder an estimate of the decoder \ac{dof} at pre-defined points of the
transmission process. The points in the process occur when a decoder
obtains a given amount of \ac{dof}. In our implementation, we define $r(k)$ as the \ac{dof}s of the decoder where it has to send the $k$-th feedback in
\eqref{eq:feedback_dofs}.
%
\begin{equation}\label{eq:feedback_dofs}
    r(k) = \floor{g\cdot \left[\frac{2^k -1}{2^k}\right]},\ k \in \Big[1, 2, \ldots,\ceil{log_2(g)}+1\Big]
\end{equation}
%
At the beginning of the transmission process, we assume that the encoder
starts with an initial coding density that has been calculated depending to
the budget. According to \eqref{eq:feedback_dofs}, a decoder sends feedback
once it has obtained (rounded down) $g/2,\ 3g/4,\ 7g/8,\ \ldots$  degrees of
freedom. The total amount of feedback for this scheme will depend on the
generation size. Still, we have an implementation that limits the amount of
feedback to be sent from the decoder in~\cite{practicalview_tsnc2015}.
